Watch Out for the Bully! Job Interference Study on Dragonfly Network

### 2016-04-09 study_workload_2.tex
remove TODO in Section 5
tweak in Section 5

Figure~\ref{fig:synwkld-network-traffic-stime} shows aggregate traffic and
saturation times for Workload~\Rmnum{2}, corresponding to
Figure~\ref{fig:wkld-network-traffic-stime} in Workload~\Rmnum{1}.  Replacing
AMG with sAMg results in greater aggregate traffic as well as more saturation
than in Workload~\Rmnum{1}, but regardless, similar patterns can be observed.
Contiguous placement with minimal routing results in load imbalances with
respect to both traffic and saturation. The addition of adaptive routing
alleviates these effects to some degree, particularly with respect to global
channel usage, while trading off saturation in global channels for saturation in
the local channels. Using random placement again shows roughly similar performance
characteristics across routing configurations, with adaptive routing helping to
balance aggregate load while increasing the aggregate traffic due to the related
indirection.

@@

We study the performance of each application individually in the same manner as
in Section~\ref{sec: workload-1 app analysis}.
Figure~\ref{fig:syn-apps-commtime} shows the communication time distributions of
the ranks of the three applications, running concurrently in Workload~\Rmnum{2}.
The ``bully" in Workload~\Rmnum{1} becomes the ``bullied" in Workload~\Rmnum{2}. 
MultiGrid and CrystalRouter are in this instance the less communication-intensive applications. 
With random placement and (progressive) adaptive routing, 
MultiGrid and CrystalRouter experience prolonged communication time, 
as shown in Figure~\ref{fig:syn-mg-commtime}, \ref{fig:syn-cr-commtime}. 
On the other hand, sAMG (Figure~\ref{fig:syn-samg-commtime}) benefits from those configurations
in a similar manner to CrystalRouter in Workload~\Rmnum{1}.
Contiguous placement coupled with minimal routing, while preventing the bully
behavior, results in poor performance for all of the applications except CrystalRouter,
which we also expect is due to a higher degree of network isolation.

@@

MultiGrid in Workload~\Rmnum{2} appears most similar to AMG in Workload~\Rmnum{1} when considering resident channel behavior, as shown in Figures~\ref{fig:syn-mg-lc-traffic} and \ref{fig:syn-mg-gc-traffic}. There is a large gap in traffic volume between the contiguous and random placement approaches, due to other applications (sAMG in particular) utilizing the same links. CrystalRouter in Workload~\Rmnum{2} additionally experiences more load on its channels using random allocation configurations, as shown in Figure~\ref{fig:syn-cr-lc-traffic} and \ref{fig:syn-cr-lc-traffic}. However, the maximal load under random allocation is closer to that seen in contiguous allocations as compared to MultiGrid.

@@

\subsection{Observations}

\NOTE{rewrote this somewhat. The observations made were essentially the same
as the previous section}

Revisiting the observations in Section~\ref{sec:workload-1-observations}, we
find that those observations are held under this separate configuration.
System-level performance is still much improved in terms of load-balancing.
sAMG, being far and away the most communication-intensive application, benefits
greatly from random placement, whereas in Workload~\Rmnum{1} AMG was effectively
penalized for being less communication-intensive. CrystalRouter, being the
comparatively less communication-intensive application in Workload~\Rmnum{2},
experiences performance regressions in Workload~\Rmnum{2} under random and adaptive
policies.

Interestingly, in both Workload~\Rmnum{1} and \Rmnum{2}, MultiGrid experiences
a more subtle performance behavior than the significant swings in performance
seen in the other applications. These behaviors persisted across multiple runs
with different random seeds. Additionally, CrystalRouter in Workload~\Rmnum{2}
sees less drastic changes in maximal load, but still experiences performance
regressions. We are continuing to work towards understanding the root causes
and implications of this behavior, for which we expect application-specific
communication patterns to be an important factor. \TODO{Is this a reasonable
observation?}

>>>>

Figure~\ref{fig:synwkld-network-traffic-stime} shows aggregate traffic and
saturation times for Workload~\Rmnum{2}, corresponding to
Figure~\ref{fig:wkld-network-traffic-stime} in Workload~\Rmnum{1}. Replacing
AMG with sAMG results in greater aggregate traffic as well as more saturation
than in Workload~\Rmnum{1}, but regardless, similar patterns can be observed.
Contiguous placement with minimal routing results in load imbalance with
respect to both traffic and saturation. The addition of adaptive routing
alleviates these effects to some degree, particularly with respect to global
channel usage, while trading off saturation in global channels for saturation in
the local channels. Using random placement again shows roughly similar performance
characteristics across routing configurations, with adaptive routing helping to
balance aggregate load while increasing the aggregate traffic due to the related
indirection.

@@

We study the performance of each application individually in the same manner as
in Section~\ref{sec: workload-1 app analysis}.
Figure~\ref{fig:syn-apps-commtime} shows the communication time distributions of
the ranks of the three applications, running concurrently in Workload~\Rmnum{2}.
The ``bully" in Workload~\Rmnum{1} becomes the ``bullied" in Workload~\Rmnum{2}. 
MultiGrid and CrystalRouter are in this instance the less communication-intensive applications. 
With random placement and (progressive) adaptive routing, 
both MultiGrid and CrystalRouter experience prolonged communication time, 
as shown in Figure~\ref{fig:syn-mg-commtime}, \ref{fig:syn-cr-commtime}. 
On the other hand, sAMG (Figure~\ref{fig:syn-samg-commtime}) benefits from those configurations
in a similar manner to CrystalRouter in Workload~\Rmnum{1}.
Contiguous placement coupled with minimal routing, while preventing the ``bully"
behavior, results in poor performance for all of the applications except CrystalRouter,
which we expect is due to a higher degree of network isolation.

@@

MultiGrid in Workload~\Rmnum{2} is similar to AMG in Workload~\Rmnum{1} when considering resident channel behavior, as shown in Figures~\ref{fig:syn-mg-lc-traffic} and \ref{fig:syn-mg-gc-traffic}. There is a large gap in traffic volume between the contiguous and random placement approaches, due to other applications (sAMG in particular) utilizing the same links. CrystalRouter in Workload~\Rmnum{2} additionally experiences more load on its channels using random allocation configurations, as shown in Figure~\ref{fig:syn-cr-lc-traffic} and \ref{fig:syn-cr-lc-traffic}. However, the maximal load under random allocation is closer to that observed in contiguous allocations as compared to MultiGrid.

@@

\subsection{Key Observations}

%\NOTE{rewrote this somewhat. The observations made were essentially the same as the previous section}

Revisiting the observations in Section~\ref{sec:workload-1-observations}, we
find that those observations are held under this separate configuration.
%System-level performance is still much improved in terms of load-balancing.
System-level performance is still much improved in terms of load-balancing with random placement.
sAMG, being far and away the most communication-intensive application in Workload~\Rmnum{2}, benefits
greatly from random placement, whereas in Workload~\Rmnum{1} AMG was effectively
penalized for being less communication-intensive. CrystalRouter, being the
comparatively less communication-intensive application in Workload~\Rmnum{2},
experiences performance regressions in Workload~\Rmnum{2} under random and adaptive
policies.

Interestingly, in both Workload~\Rmnum{1} and \Rmnum{2}, MultiGrid experiences
a more subtle performance variatioin than the significant swings in performance
observed in the other applications. These behaviors persisted across multiple runs
with different random seeds. Additionally, CrystalRouter in Workload~\Rmnum{2}
exhibits less drastic changes in maximal load, but still experiences performance
regressions. We are continuing to work towards understanding the root causes
and implications of this behavior, for which we expect application-specific
communication patterns to be an important factor. 
%\TODO{Is this a reasonable observation? Yes, it is.}
